---
title: "p8105 Homework 6"
output: github_document
date: "2024-12-03"
---
Name [UNI]: Xi Peng [xp2213]

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(broom)
library(purrr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(924)
```

# Question 1: 2017 Central Park weather data
```{r}
# Dataset download and processes.
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

# Define a bootstrap sampling function.
boot_sample = function(df) {
  sample_frac(df, replace = TRUE)
}

# Bootstrap resampling.
boot_straps_result = tibble(strap_number = 1:5000) |> 
  mutate(
    strap_sample = map(strap_number, ~ boot_sample(weather_df)),
    models = map(strap_sample, ~ lm(tmax ~ tmin, data = .x)),
    results = map(models, tidy),                         
    r_squared = map_dbl(models, ~ glance(.x)$r.squared), 
    log_beta_product = map_dbl(models, ~ {       
      coefs = tidy(.x)  
      beta_0 = coefs |> filter(term == "(Intercept)") |> pull(estimate)
      beta_1 = coefs |> filter(term == "tmin") |> pull(estimate)
      log(abs(beta_0 * beta_1))
    })
  )

boot_straps_result
```
<br>
Plot the distribution of the estimates of the two quantities.
```{r}
boot_straps_plot = 
  boot_straps_result |>
  select(r_squared, log_beta_product) |> 
  pivot_longer(cols = everything(), names_to = "metric", values_to = "value") |> 
  ggplot(aes(x = value)) +
  geom_density(alpha = 0.3) +
  facet_wrap(~metric, scales = "free") +
  labs(
    title = "Bootstrap Distributions of R^2 and log(β0 * β1)",
    x = "Value",
    y = "Density"
  )

boot_straps_plot
```

Both distributions show clear, bell-shaped curves with minimal skewness, indicating that these statistics are stable and reliable across different bootstrap samples.The plot indicates a strong linear relationship between tmax and tmin in the weather data, as reflected by the consistently high r_squared values and stable log_beta_product values across 5000 bootstrap samples. The narrow distribution of r_squared values demonstrates that the model explains a substantial proportion of the variance in tmax, while the consistency of the log_beta_product values indicates minimal variability in the relationship between the coefficients. The large sample size of 5000 ensures that the distributions capture the true underlying pattern with minimal variation across resampled datasets.

<br>
95% confidence interval for the two quantities.
```{r}
CI_result = boot_straps_result |> 
  summarize(
    r_squared_CI = list(quantile(r_squared, c(0.025,0.975))),
    log_beta_product_CI = list(quantile(log_beta_product, c(0.025, 0.975)))
  ) 

knitr::kable(CI_result)
```








